---
title: "data_viz_morning"
author: "Juliet"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(janitor)
library(tidyverse)

# for data viz:
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(beeswarm)
#this one is for data:
library(gapminder)
#go-to for working with spatial data in R, propietary software:
library(sf)

```

## get the lizards data

```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))

#to check working directory, in console type getwd() or here::here()
```
# Make a (bad) historgram of lizard weights

```{r}
lizard_weights <- ggplot(data = lizards, aes(x = weight)) +
  geom_histogram(fill = "orange",
                 color = "cyan",
                 size = 0.8,
                 linetype = "dotted")

# AKA

#lizard_weights <- ggplot() +
#  geom_histogram(data = lizards, aes(x = weight))

#this second way gives you flexibility to overlay 2 plots from 2 diff datasets

# color = points and lines
# fill = bars and such
# size = changes line width, can be whole numbers like 3 or decimal to make smaller like 0.2

lizard_weights
```

## Make a scatterplot, total_length = x, weight = y

```{r}
lizards_scatter <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(shape = 22,
             fill = "yellow",
             color = "red",
             size = 2,
             alpha = 0.5)

# shape numbers represent different shapes for the points, 10-22 is a good range but it geos higher to random numbers
# some point styles only have color, some have color and fill

lizards_scatter
```

## Make a bubble plot where color changes based on common name and size varies based on total length

```{r}
lizards_bubble <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name,
                  size = total_length), 
             shape = 22,
             fill = "black") +
  theme_get()

lizards_bubble
```
## Facet

```{r}
#now consider facet_wrap vs facet_grid (creates a grid based on 2 cat. var, and populates each intersection with a diff graph), facet_wrap interprets the number of graphs you want based on the groups it recognizes in the col you give it with ~col
# with facet wrap the argument ncol = # means you tell R how many graphs to put into each row

lizards_facet_wrap <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  theme_get() +
  facet_wrap(~common_name, ncol = 10)

lizards_facet_wrap

#---------------------------
#facet_grid(rows ~ col)

lizards_facet_grid <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  facet_grid(sex ~ tail)

lizards_facet_grid
```
# Find lizard counts by common name

```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

lizard_counts

# AKA

lizard_counts <- lizards %>% dplyr::count(common_name)

lizard_counts

lizard_counts_cn_tail <- lizards %>% count(common_name, tail)
lizard_counts_cn_tail

# also consider tally() but not as good


```

```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

lizard_counts

#you can specify x and y and not put them in a particular order
# here, we want the y axis to have the names beause they are long and it is best to have them readable on the y-axis so they arent crowded

ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, count), x = count)) +
  geom_col(aes(fill = common_name, show.legend = FALSE)) +
  labs(x = "Lizard Count")

# can also put coord_flip() instead of specifying y = common_names and x = count in aes(), but this is less direct and more confusing
# when you convert to a factor, ggplot's default is alphabetical, but if you want a particular order you need to make it a factor and assign the factor levels based on the value assigned in another variable
# this package is "forcats" in the tidyverse
```

Try converting common_name to an ordered factor outside of ggplot

```{r}
change_common_name <- lizard_counts %>% 
  mutate(common_name = fct_reorder(common_name, count))
         
class(change_common_name$common_name)
levels(change_common_name$common_name)
```

# Axis sclae breaks, limits, and labels


Scatterplot: total_length vs weight (y)
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  scale_x_continuous(breaks = c(0, 50, 500),
                     limits = c(0, 500),
                     expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0))

# want an ending tick mark that includes all data points, but in the original but there is an outlier
# there are too many breaks in the gridline, too. when fixing this, specify the axis and type of variable
# scale_ brings up tonnnnnns of options with options for cutomizations
# breaks = c() or seq()
```

## transform the date col to class Date, then find counts of obs by date

```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  count(date)

#converting this to a date allows R to make sense of the date col, otherwise the characters are just jumbled
  
class(lizard_counts$date)
```

## making a line plot of date (x) and count (y)

```{r}
ggplot(data = lizard_counts, aes(x = date, y = n)) +
  geom_line() +
  scale_x_date(date_breaks = "3 years",
               date_labels = "%Y")

# R makes an increment on the x axis for time, if you didnt convert the class to date this would not have been executed right
# %y means 2 -digit date and %Y means 4 -digit date
```
## color_gradient()

```{r}
#make your own gradient with scale-color_gradient(low =, high=)

lizards_scatter <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradient(low = "red", high = "navy")

lizards_scatter

#scale_color_gradientn

lizards_scatter <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradientn(colors = c("orange", "red", "purple", "black"))

lizards_scatter

# can have meaningful color breaks like if you are showing thresholds for air quality, use color steps

lizards_scatter <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_stepsn(colors = c("green", "blue", "purple"),
                     breaks = c(0, 20, 40, 50))

lizards_scatter
```
### discrete levels of a variable : update a color scheme using a palette in paleteer
Make a horizontal box plot with common name on y, total length on x, with fill color changing based on common name
```{r}

lizards_fct <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))

# check the order it made with levels()

levels(lizards_fct$common_name)

ggplot(data = lizards, aes(x = total_length, y = common_name)) +
  geom_boxplot(aes(fill = common_name),
               legend = FALSE) +
  scale_fill_paletteer_d(palette = "tvthemes::EarthKingdom")
```

## Themes

```{r}
lizards_scatter <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
    panel.background = element_rect(fill = "yellow",
                                    color = "purple",
                                    size = 10),
    panel.grid.major.x = element_line(color = "red"),
    panel.grid.major = element_line(color = "blue"),
    panel.grid.minor.y = element_line(color = "orange")
  )
#the most specific level trumps other vague levels when you are working in theme
#panel.grid = element_blank() gets rid of the gridlines
lizards_scatter

```

```{r}
lizards_scatter <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
    panel.grid = element_blank(),
    plot.background = element_rect(fill = "cyan"),
    axis.text.x = element_text(color = "orange"),
    axis.title = element_text(color = "yellow", size = 19)
  )

lizards_scatter

```














